Increasing Performance via Gamification in a Volunteer-Based Evolutionary Computation System

### 2017-11-10 2017-EUROPAR/Gamification/paper.tex
Responding to one of the reviewers
Saying that we are more interested in the real amount of time to see
if it is worth the while. Please @mariosky check it out.

The world wide web is an increasingly reliable and high-performance
operating system for running distributed applications. Besides the
maturity of HTTP, the underlying protocol itself, there are other
factors: the JavaScript virtual machine embedded in every browser
and the REST interface, providing a {\em de facto} standard interface for
communication. A distributed system based on the browser can be
created easily using these mechanisms, which have lately been enhanced
by WebSockets for lower latency; you can gather users by just
announcing the URL. This approach for creating distributed experiments is called {\em
  volunteer}, {\em cycle-scavenging}, or {\em opportunistic} computing
\cite{sarmenta2001volunteer} and it dates back, in different shapes
and underlying mechanisms, to the origin of the web
\cite{david-seti:home}.

We are mainly interested in evolutionary algorithms \cite{jj-ppsn98},, population-based search and optimization methods inspired by
the theory of Evolution and its molecular basis. The fact that they are population-based makes them suitable for a straightforward distribution of work to different clients, and since they have an asynchronous nature, there is not a considerable impact on their performance. All these properties make them ideal candidates for volunteer computing setups such as the one presented in this paper.

There are several open issues in the research regarding the
application of volunteer computing techniques to the development of
distributed evolutionary algorithms of which the most important is
approaching the problem as a socio-technical system
\cite{vespignani2009predicting,merelo2015designing}. which integrates
user decisions and behavioral patterns in the model; this
includes trying to optimize the number of users in a particular
experiment. The challenge is to design a system that, whatever the
number of available users to perform the experiment, it is able to
maximize their contribution to the evolutionary algorithm. One obvious
way of achieving that is trying to find as many users as possible. But
another approach, and the one that we will be following in this paper, is
to delve into the socio-technical nature of the system, using
gamification techniques to improve the amount of time every user will
lend to the system. Gamification has to include some way of user authentication,
which is a hurdle that might take users away; however, the open design
of the system we use means that we can actually mix anonymous-non
gamified and authenticated-gamified users, with only the small overhead
of adding the gamification mechanism.

There is a trade-off between ease of use, allowing users to
participate by just visiting a web page, and personalization, making
registration compulsory and subsequently authenticate in order to
participate. In a sociotechnical system such as this one, every
decision the user has to take has an influence on the overall
performance, and registration could be seen as a hurdle to volunteers
trying to participate in an ephemeral experiment, so it is a challenge
to analyze what actually happens in this case; this is one of the main
questions addressed in this paper.

>>>>

The world wide web is an increasingly reliable and high-performance
operating system for running distributed applications. Besides the
maturity of HTTP, the underlying protocol itself, there are other
factors: the JavaScript virtual machine embedded in every browser
and the REST interface, providing a {\em de facto} standard interface for
communication. A distributed system based on the browser can be
created easily using these mechanisms, which have lately been enhanced
by WebSockets for lower latency; you can gather users by just
announcing the URL. This approach for creating distributed experiments is called {\em
  volunteer}, {\em cycle-scavenging}, or {\em opportunistic} computing
\cite{sarmenta2001volunteer} and it dates back, in different shapes
and underlying mechanisms, to the origin of the Internet
\cite{david-seti:home}.

In this context, we are mainly interested in evolutionary algorithms
\cite{milani2004online,sherry2012flex,jj-ppsn98}%,
% population-based search and optimization methods inspired by
% the theory of Evolution and its molecular basis.
. The fact that they are population-based makes them suitable for a
straightforward distribution of work to different clients, and since
they have an asynchronous nature, there is no considerable impact on
their performance, and might even be beneficial for a number of
reasons, including increased diversity
\cite{cantu-paz:migration-policies}. All these properties make them
ideal candidates for volunteer computing setups such as the one
presented in this paper, at least if some minimal changes are made to
the usual sequential or parallel algorithms that rely on a static
setup.

But even if adaptation is straightforward, there are several open issues in the research regarding the
application of volunteer computing techniques to the development of
distributed evolutionary algorithms of which the most important is
approaching the problem as a socio-technical system
\cite{vespignani2009predicting,merelo2015designing}, which integrates
user decisions and behavioral patterns in the model; this
includes trying to optimize the number of users or the overall time
they contribute in a particular
experiment. The challenge is to design a system that, whatever the
number of users available and willing to perform the experiment, it is able to
maximize their contribution to the evolutionary algorithm. One obvious
way of achieving that is trying to find as many users as possible. But
another approach, and the one that we will be following in this paper, is
to delve into the socio-technical nature of the system, using
gamification techniques to try and improve the amount of time every user will
lend to the system. However, there is a tradeoff with gamification,
since users have to be identified in some way to participate, and thus
it has to include some way of user authentication,
which is a hurdle that might take users away; however, the open design
of the system we use means that we can actually mix anonymous-non
gamified and authenticated-gamified users, with only the small overhead
of adding the gamification mechanism.

This trade-off between ease of use, allowing users to
participate by just visiting a web page, and personalization, making
registration compulsory and subsequently authenticate in order to
participate, is what we will analyze in this paper. In a sociotechnical system such as this one, every
decision the user has to take has an influence on the overall
performance, and registration could be seen as a hurdle to volunteers
trying to participate in an ephemeral experiment, so it is a challenge
to analyze what actually happens in this case; quite clearly, the
inclusion of gamification techniques has an impact on page loading and
complexity of server-side programming; if eigher the number of users
or the amount of time they contribute to the experiment is not enough,
it would be better to revert to the purely anonymous system.
